---
title: "Portfolio-Part-1-Development-and-Planning"
format: html
echo: false
---


Data Description
Identify your data source.
> My dataset is about the TV show The Simpsons. The dataset is from a tidytuesday submission, where the data was extracted from imdb using Kaggle.  The data is froom approximately 600 simpsons episodes from 2010 to 2016.

Describe your data, including variables and data types.
> The dataset is split in four csv files, with each file have varying variable types. The simpson_characters csv has the id, name, normalized name, and gender of each character from the recorded episodes.  The id is a double, while the rest of the variables are characters. The simpsons_episodes csv has id, image_url, imdb_rating, imdb_votes, number_in_season, number_in_series, original_air_date, original_air_year, production_code, season, title, us_viewers_in_millions, video_url, and views as the variables.  There are doubles and characters for the variable types in this dataset.

Identify the research questions you want to answer.
> The potential research questions I would like to answer would be 
1. Do us viewers in millions increase over progession in episodes of a season.
2. Do seasons of the Simpsons in the 1990s have a higher imdb rating on average comapred to Seasons from the 2000s

Data Visualization
What do you want your final visualizations to look like?

> I want scatterplots to show if there is clear representation of viewer retention for episodes as a season of the simpsons progresses.

What do you want to highlight on your final visualizations in order to answer your research questions? How do you plan to do that?
What is missing from your data or would need to change in your data to create these visualizations?


Data Cleaning
The answer to at least three of these questions should be “YES” for the data to meet the necessary standards to demonstrate your cleaning. Your data source should not be an already perfectly prepared data set.

Do you need to reformat any variables into different types (e.g. factors, time, dates, strings)? Or remove information from variable values?
Do you need to deal with any missing data, especially missing data coded other than NA?
Do you need to filter your data? How?
> Yes, filtering data by season and year

Do you need to create any new variables? What variables? How?
Do you need to add new data (join) to your data? What data? How?
Do you need pivot your data in any way? Why? How?
Do you need to summarize any of the variables? Which ones? How?
> Yes, I need to summarize the imdb score

What other aspects of your data need to be “fixed” in order to make your data visualizations?
Most will answer yes to the following for making your programming more efficient using select(), but you should have three other yeses above.

Are there any variables you can exclude from your data?
> Yes, the production code variable is not needed


```{r}
library(tidyverse)
library(janitor)
```



```{r}
simpsons_characters <- read.csv("simpsons-dataset/simpsons_characters.csv")
simpsons_episodes <- read.csv("simpsons-dataset/simpsons_episodes.csv")

simpsons_episodes_clean <- clean_names(simpsons_episodes)



simpsons_episodes_clean |>
  filter(season == 1) |>
  ggplot() +
  geom_point(mapping = aes(x = number_in_season, y = us_viewers_in_millions),
             color = "blue") +
  xlab("Number of episode in season") +
  ylab("U.S. Viewers in millions")
           

simpsons_episodes_clean |>
  filter(season == 2) |>
  ggplot() +
  geom_point(mapping = aes(x = number_in_season, y = us_viewers_in_millions), 
           color = "green") +
           xlab("Number of episode in season") +
           ylab("U.S. Viewers in Millions")


simpsons_episodes_clean |>
  filter(season == 3) |>
  ggplot() +
  geom_point(mapping = aes(x = number_in_season, y = us_viewers_in_millions), 
           color = "red") +
           xlab("Number of episode in season") +
           ylab("U.S. Viewers in Millions")

  
```
```{r}
simpsons_1990 <- mutate(simpsons_episodes_clean) |>
  filter(original_air_year < 2000)





```


